---
title: "Covid 19 dashboard for Sri Lanka"
author: "Pushpakanthie Wijekoon"
date: "Last update: `r format(Sys.time(), '%B %d, %Y')`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    #source_code: embed
    vertical_layout: scroll
---


```{r setup, include=FALSE}
#------------------ Packages ------------------
library(flexdashboard)
# install.packages("devtools")
# devtools::install_github("RamiKrispin/coronavirus", force = TRUE)
library(coronavirus)
data(coronavirus)
# View(coronavirus)
# max(coronavirus$date)

`%>%` <- magrittr::`%>%`
#------------------ Parameters ------------------
# Set colors
# https://www.w3.org/TR/css-color-3/#svg-color
confirmed_color <- "#800080"
active_color <- "#1f77b4"
recovered_color <- "#808000"
death_color <- "#CB4154"
#------------------ Data ------------------

df <- read.csv("https://raw.githubusercontent.com/RamiKrispin/coronavirus/master/csv/coronavirus.csv", stringsAsFactors = FALSE) 
df %>%
   # dplyr::filter(date == max(date)) %>%
  dplyr::filter(country == "Sri Lanka") %>%
  dplyr::group_by(country, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  # dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(-confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(country == "United Arab Emirates", "UAE", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

df_daily <- df %>%
  dplyr::filter(country == "Sri Lanka") %>%
  dplyr::group_by(date, type) %>%
  dplyr::summarise(total = sum(cases, na.rm = TRUE)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  dplyr::arrange(date) %>%
  dplyr::ungroup() %>%
  #dplyr::mutate(active = confirmed - death - recovered) %>%
  dplyr::mutate(active = confirmed - death) %>%
  dplyr::mutate(
    confirmed_cum = cumsum(confirmed),
    death_cum = cumsum(death),
    # recovered_cum = cumsum(recovered),
    active_cum = cumsum(active)
  )

df1 <- df  %>% dplyr::filter(date == max(date))

```
  
  
```{r}
df_tree <- df %>%
  dplyr::group_by(country, type) %>%
  dplyr::summarise(total = sum(cases), .groups = "drop") %>%
  dplyr::mutate(type = ifelse(type == "confirmed", "Confirmed", type),
                type = ifelse(type == "recovered", "Recovered", type),
                type = ifelse(type == "death", "Death", type)) %>%
  tidyr::pivot_wider(names_from = type, values_from = total) %>%
  dplyr::mutate(Active = Confirmed - Death - Recovered) %>%
  tidyr::pivot_longer(cols = -country, names_to = "type", values_to = "total")

df_world <- df_tree %>%
  dplyr::group_by(type) %>%
  dplyr::summarise(total = sum(total), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = type, values_from = total)

names(df_world) <- tolower(names(df_world))
```
  

Row {data-width=400}
-----------------------------------------------------------------------
  
### confirmed {.value-box}      

```{r}
valueBox(
  value = paste(format(sum(df_daily$confirmed), big.mark = ","), "", sep = " "),
  caption = "Total confirmed cases",
  icon = "fas fa-user-md",
  color = confirmed_color
)
```
  
### recovered {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df_daily$recovered, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df_daily$recovered, na.rm = TRUE) / sum(df_daily$confirmed), 2),
    "%)",
    sep = ""
  ),
  caption = "Recovered Cases (recovered rate)",
  icon = "fas fa-heartbeat",
  color = recovered_color
)
```


### death {.value-box}         

```{r}
valueBox(
  value = paste(format(sum(df_daily$death, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df_daily$death, na.rm = TRUE)/ sum(df_daily$confirmed), 2),
    "%)",
    sep = ""
  ),
  caption = "Death cases (death rate)",
  icon = "fas fa-heart-broken",
  color = death_color
)
```

   
Row {.tabset}  
-----------------------------------------------------------------------
  

### **Daily cumulative cases**    

```{r}
plotly::plot_ly(data = df_daily, height=400, width=900)%>%
  plotly::add_trace(
    x = ~date,
    # y = ~active_cum,
    y = ~confirmed_cum,
    type = "scatter",
    mode = "lines+markers",
    # name = "Active",
    name = "Confirmed",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~death_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Death",
    line = list(color = death_color),
    marker = list(color = death_color)
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-01-27"),
    y = 1,
    text = paste("First case"),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -10,
    ay = -90
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-28"),
    y = 3,
    text = paste("First death"),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = 90,
    ay = -90
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-20"),
    y = 14,
    text = paste(
      "Lockdown styled curfew"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -20,
    ay = -90
  ) %>%
  plotly::layout(
    title = "",
    yaxis = list(title = "Cumulative number of cases",titlefont = list(size = 9), tickfont = list(size = 9)),
    xaxis = list(title = "Date",titlefont = list(size = 9), tickfont = list(size = 9)),
    legend = list(x = 0.4, y = 1.5, font = list(size = 10)),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 5,
      t = 5
      #pad = 2
      )
  )
```


### **Daily confirmed cases**  

```{r}
daily_confirmed <-df %>%
  dplyr::filter(type == "confirmed") %>%
  dplyr::filter(date >= "2020-01-27") %>%
  dplyr::mutate(country = country) %>%
  dplyr::group_by(date, country) %>%
  dplyr::summarise(total = sum(cases)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = country, values_from = total)
#names(daily_confirmed)[names(daily_confirmed) == 'Sri Lanka'] <- 'sl'

#----------------------------------------
# Plotting the data

daily_confirmed %>%
  plotly::plot_ly(height=400, width=900) %>%
  plotly::add_trace(
    x = ~date,
    y = ~`Sri Lanka`,
    type = "scatter",
    mode = "lines+markers",
    name = "Sri Lanka"
  ) %>%
  plotly::add_trace(
     x = ~date,
     y = ~Maldives,
     type = "scatter",
     mode = "lines+markers",
     name = "Maldives"
   ) %>%
   plotly::add_trace(
     x = ~date,
     y = ~Bhutan,
     type = "scatter",
     mode = "lines+markers",
     name = "Bhutan"
   ) %>%
  #plotly::add_trace(
    #x = ~date,
    #y = ~Burma,
    #type = "scatter",
    #mode = "lines+markers",
    #name = "Burma"
  #) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Vietnam,
    type = "scatter",
    mode = "lines+markers",
    name = "Vietnam"
  ) %>%
  #plotly::add_trace(
    #x = ~date,
    #y = ~Singapore,
    #type = "scatter",
    #mode = "lines+markers",
    #name = "Singapore"
  #) %>%
  plotly::layout(
    title = "",
    legend = list(x = 0.1, y = 0.98),
    yaxis = list(title = "Daily confirmed cases",titlefont = list(size = 9), tickfont = list(size = 9)),
    xaxis = list(title = "Date", titlefont = list(size = 9), tickfont = list(size = 9)),
    # paper_bgcolor = "black",
    # plot_bgcolor = "black",
    # font = list(color = 'white'),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 5,
      t = 5
      #pad = 2
    )
  )
```


### **Recovery and Death rates**  


```{r}
df_rates <- df_tree %>%
  dplyr::filter(type != "Active") %>%
  tidyr::pivot_wider(names_from = "type", values_from = "total") %>%
  dplyr::mutate(recovery_rate = Recovered / Confirmed,
    death_rate = Death / Confirmed) 


bar_chart <- function(label, width = "100%", height = "14px", fill = "#00bfc4", background = NULL) {
  bar <- htmltools::div(style = list(background = fill, width = width, height = height))
  chart <- htmltools::div(style = list(flexGrow = 1, marginLeft = "6px", background = background), bar)
  htmltools::div(style = list(display = "flex", alignItems = "center"), label, chart)
}

tbl <- reactable::reactable(df_rates,
                     searchable = TRUE,
                     pagination = FALSE,
                     highlight = TRUE,
                     height = 400,
                     sortable = TRUE,
                     borderless = TRUE,
                     defaultPageSize = nrow(df_rates),
                      defaultSortOrder = "desc",
                     defaultSorted = "Confirmed",
                     columns = list(
                     country = reactable::colDef(name = "Country", minWidth = 50, maxWidth = 100),
                       Confirmed = reactable::colDef(name = "Confirmed",  minWidth = 50, maxWidth = 120, defaultSortOrder = "desc"),
                       Recovered = reactable::colDef(name = "Recovered",  minWidth = 50, maxWidth = 100),
                       Death = reactable::colDef(name = "Death",  minWidth = 50, maxWidth = 100),
                       recovery_rate = reactable::colDef(name = "Recovery Rate",  minWidth = 50, maxWidth = 200,
                                                        defaultSortOrder = "desc",
                                                      cell = function(value) {
                                                        # Format as percentages with 1 decimal place
                                                        value <- paste0(format(round(value * 100, 2), nsmall = 1), "%")
                                                        bar_chart(value, width = value, fill = "green", background = "#e1e1e1")
                                                      },
                       align = "left"),
                       death_rate = reactable::colDef(name = "Death Rate",  
                                                      minWidth = 50, maxWidth = 200,
                                                      defaultSortOrder = "desc",
                                                      cell = function(value) {
                                                        # Format as percentages with 1 decimal place
                                                        value <- paste0(format(round(value * 100, 2), nsmall = 1), "%")
                                                        bar_chart(value, width = value, fill = "red", background = "#e1e1e1")
                                                      },
                       align = "left"))
)

library(htmltools)
htmltools::div(class = "standings",
  htmltools::div(class = "title",
  htmltools::h4(strong("Total Number of Covid19 Cases by Country")),
    h6("Click on the columns names to resort the table"
  )),
  tbl,
  paste("Data last updated on", max(df$date))
)

```
  
### **Total Cases comparison**    

```{r}
df_t2 <- df_tree %>%
  dplyr::filter(country == "Sri Lanka" |
    country == "Singapore" |
    country == "Afghanistan" |
    country == "Bangladesh" |
    country == "Bhutan" |
    country == "Burma" |
    country == "Cambodia" | 
    country == "Indonesia" |
    country == "Malaysia" |
    country == "Maldives" |
    country == "Nepal" |   
    country == "Pakistan" |
    country == "Philippines" |
    country == "Taiwan" |
    country == "Thailand" |
    country == "Vietnam")

plotly::plot_ly(height=300, width=900,
  data = df_t2 %>% dplyr::filter(type == "Confirmed"),
  type= "treemap",
  values = ~total,
  labels= ~ country,
  parents=  ~type,
  domain = list(column=0),
  name = "Confirmed",
  textinfo="label+value+percent parent"
) %>%  
 plotly::add_trace(
    data = df_t2 %>% dplyr::filter(type == "Recovered"),
    type= "treemap",
    values = ~total,
    labels= ~ country,
    parents=  ~type,
    domain = list(column=1),
    name = "Recovered",
    textinfo="label+value+percent parent"
  ) %>%
  plotly::add_trace(
    data = df_t2 %>% dplyr::filter(type == "Death"),
    type= "treemap",
    values = ~total,
    labels= ~ country,
    parents=  ~type,
    domain = list(column=2),
    name = "Death",
    textinfo="label+value+percent parent"
  ) %>%
  plotly::layout(grid=list(columns=3, rows=1))  

 
```


### **Dashboard Info**  


**The Covid 19 Dashboard for Sri Lanka**

This [Covid 19 dashboard: the case of Sri Lanka](https://www.antoinesoetewey.com/files/coronavirus-dashboard.html) provides an overview of the 2019 Novel Coronavirus COVID-19 (2019-nCoV) epidemic for Sri Lanka. This dashboard is developed using the [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) framework in R/R Studio, and by referring the following two dashboards:  
1. [dashboard](https://ramikrispin.github.io/coronavirus_dashboard/){target="_blank"} developed by Rami Krispin, and  
2. [dashboard](https://www.antoinesoetewey.com/files/coronavirus-dashboard.html) developed by Antoine Soetewey.  

**Data**

This data set contains the following fields:  
* ```date``` - The date of the summary  
* ```province``` - The province or state, when applicable  
* ```country``` - The country or region name  
* ```lat``` - Latitude point  
* ```long``` - Longitude point  
* ```type``` - the type of case (i.e., confirmed, death)  
* ```cases``` - the number of daily cases (corresponding to the case type)  

Download the data set from  [`{coronavirus}`](https://github.com/RamiKrispin/coronavirus){target="_blank"} R package. Make sure to download the development version of the package to have the latest data:

```
install.packages("devtools")
devtools::install_github("RamiKrispin/coronavirus")
```

The data and dashboard are refreshed on a daily basis.

The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [repository](https://github.com/RamiKrispin/coronavirus-csv){target="_blank"}.  


**Information**  
More information about this dashboard can be found in this [article](https://www.statsandr.com/blog/how-to-create-a-simple-coronavirus-dashboard-specific-to-your-country-in-r/).  


**Update**  
The dashboard has been updated on `r format(Sys.time(), "%A %B %d, %Y")`.




